---
title: "R: Assigment 1"
author: "Takudzwa Machida 202002075"
date: "17/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. BASICS

## 1.1. Variable in R

#### What is the value of " B " after running these command? [2]

```{r}
a <- 5.0
c <- 10
t <- 12
B <- a+c/t
```

* The value of B is `r round(B,2)`

#### What type of vector do you get in each of the following? [4]

```{r}
vec_1 <- c(44, 5, 0, "k")
vec_2 <- c(1, 2, 3, FALSE)
vec_3 <- c("a", "b", "c", TRUE)
vec_4 <- c(1, 2, 3, "4")
```
* vec_1 we get a "character"

* vec_2 we get a "numeric"

* vec_3 we get a "character"

* vec_4 we get a "character"

## 1.2. Conditional subsetting

### Consider the following vector:

```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)
```


#### 1.2.1. Create a vector vec_1 with no missing values using conditional subsetting and explain your code [3]
```{r}
x <- is.na(vec) 
vec_1<-vec[!x] 
print(vec_1)  
```
* Contents of the vector with no missing values:
`r vec_1`

Explanation: 

- the code "is.na(vec)" searches positions in the vector where there is NA and prints True and if NA is not there  prints False.

- '!x' are the positions in the vector where we dont have NA and vec[!x] print are the known values in the vector.

- 'print(vec_1)' displays the vector with no missing terms.



#### 1.2.2. Create a vector of indexes of missing values using conditional subsetting and explain your code [3]
```{r}

y<-vec[is.na(vec)] 
print(y)   
```
Explanation:

- "is.na(vec)" are the positions where missing value (NA) are.

- "vec[is.na(vec)]" is assigned to y and if y is called the missing values NA are displayed.


#### 1.2.3. Write a code to determine how many missing values are in vec. [2]
```{r}
y<-length(vec[is.na(vec)])
print (y)

```
  - there are `r y` missing values in the vector.
  
  
#### 1.2.4. Write a R code to create a data frame which contains details of 3 students as follows: [4]
    Name,
    Surname,
    Age,
    Gender.
```{r}
name <- c("Abiola","Mufaro","Paida")
surname <- c("Machida","Vandirai","Andonate")
Age <- c(23,25,20)
Gender <- c("Female","Male","Female")
df <- data.frame(name,surname,Age,Gender)
print(df)
```


## 2. Functions in R

2.1. Create a function which returns a mean of an input vector and explain your code [4]
```{r}
The_Mean <- function(vector){
  z <- mean(vector,na.rm = TRUE)
  return(z)
}

trial_vector <- c(1,2,3,NA,5)
The_Mean(trial_vector)
```
Explanation:

- "The_Mean" is the name to the function i have created which i will call when i want to calculate the mean of the input vector.

- l have also passed one parameter in the function which represece the input vector.

- the code " z <- mean(vector,na.rm = TRUE)" looks in the vector and if there are any missing values, the are ignored and the mean if calculated


#### 2.2. Create a function which take no inputs , it reads a Titanic .csv file and returns its content and explain your code. [3] 
```{r}
library(tidyverse)
df<- function(){     #giving my function a name df
  x <- tibble(read.csv("train.csv")) # this line of code reads the csv file stored in computer named train and assigns to x
  return(x)   #after reading the csv file this line allows to output the information in the csv file
}
df()  #displays the info of the csv file
```


## 3. Conditionals and Loops

#### 3.1. Write a function that take an input below and returns a transformed vector below into 1 when its element is male and 2 when its female using ifelse. [8]
```{r}
g <- c("MALE","FEMALE","FEMALE","MALE","FEMALE","MALE","MALE","FEMALE", "MALE","MALE","FEMALE","MALE","MALE",
"FEMALE","MALE","MALE","MALE")

Gender <- function(){
  y <- ifelse(g=="MALE",1,2)
  return(y)
}
Gender()
```
*The transfomed vector where 1 is Male and 2 is Female:

`r Gender()`

Eplanation:

- "Gender" is the name i have given to the function which uses the converts male to 1 and female to 2.

- The "ifelse" statement goes through vector 'g' and if element is MALE it returns  1 and if not it returns 2.


#### 3.2. Generate x from normal distribution and sample with replacement "M" and "F" for male and female respectively by running the code below.

```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
```

3.2.1. Use the "cbind " function in R to create one dataset and assign it to df. [2]
```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
df <- cbind(x,y)

print(head(df))  # the first 6 elements 
```


3.2.2. Convert df from 3.2.1 into a data.frame [2]
```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
#df from 3.2.1
df <- data.frame(df)

#print(head(df))  # first 6
```


3.2.3. Write a code using tidyverse to compute a number of males ("M") and females ("F") in 3.2.1 [3]
```{r}
library(tidyverse)
df %>% group_by(y) %>% count()
```


3.2.4. Create a bar plot using ggplot to visualize your results in 3.2.3 [5]
```{r}
library(ggplot2)
df %>% group_by(y) %>% count %>% 
  ggplot(aes(x=y,y=n,))+
  geom_col()+
  labs(y="total number",x="Gender",title="Plot of type of Gender against the total Number")+
  theme(plot.title=element_text(hjust = 0.5))
  

```


3.2.5. Use mutate function to add sum of x per "M" and "F" and plot the results using sum of x per "M" and "F". [5]
```{r}
df %>% group_by(y) %>% count() %>% 
  mutate(new_col=sum(n)) %>% 
  ggplot(aes(x=y,y=new_col))+
  geom_col()+
  labs(y="Sum per M and F",x="Gender",title="Plot of results using sum of x per M and F")+
  theme(plot.title=element_text(hjust = 0.5))
  


```

